Indexed Containers

### 2011-05-10 journal/paper.tex
some rejigging, added some works to bckground.

To model this we move to the notion of an inductive
family in Type Theory. A family is a type indexed by another already
given type. The first example is the family of finite sets \ensuremath{\Conid{Fin}} which
assigns to any natural number \ensuremath{\Varid{n}} a set \ensuremath{\Conid{Fin}\;\Varid{n}} which has exactly
\ensuremath{\Varid{n}}-elements. \ensuremath{\Conid{Fin}} can be used where in conventional reasoning we
assume any finite set, e.g. when dealing with a finite address apce or
a finite set of variables. The inductive definition of \ensuremath{\Conid{Fin}} refines
the type of natural numbers:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\Conid{Fin}\;\in\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{zero}\;{}\<[9]%
\>[9]{}\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;{}\<[30]%
\>[30]{}\rightarrow\;\Conid{Fin}\;(\Varid{suc}\;\Varid{n}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{suc}\;{}\<[9]%
\>[9]{}\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;(\Varid{i}\;\in\;\Conid{Fin}\;\Varid{n})\;{}\<[30]%
\>[30]{}\rightarrow\;\Conid{Fin}\;(\Varid{suc}\;\Varid{n}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

@@

\noindent
Inductive families are the backbone of
dependently typed programming as present in Epigram or
Agda~\cite{Agda}. Coq also supports the definition of inductive families
but programming with them is rather hard --- a situation which has been
improved by the new \texttt{Program} tactic~\cite{sozeau}. 
More recently, the implementation of Generalized Algebraic Datatypes 
(GADTs)~\cite{Hinze:GADT} 
allows \ensuremath{\Conid{Fin}} and \ensuremath{\Conid{Lam}} to be encoded in Haskell:
\begin{tabbing}\tt
~data~Fin~a~where~\\
\tt ~~~FZero~\char58{}\char58{}~Fin~\char40{}Maybe~a\char41{}\\
\tt ~~~FSucc~\char58{}\char58{}~Fin~a~\char45{}\char62{}~Fin~\char40{}Maybe~a\char41{}\\
\tt ~\\
\tt ~data~Lam~a~where~\\
\tt ~~~Var~\char58{}\char58{}~Fin~a~\char45{}\char62{}~Lam~a\\
\tt ~~~App~\char58{}\char58{}~Lam~a~\char45{}\char62{}~Lam~a~\char45{}\char62{}~Lam~a\\
\tt ~~~Abs~\char58{}\char58{}~Lam~\char40{}Maybe~a\char41{}~\char45{}\char62{}~Lam~a
\end{tabbing}
Here \texttt{Fin} and \texttt{Lam} are indexed by types instead of
natural numbers; The type constructor \texttt{Maybe} serves as a type level
copy of the $\succ$ constructor for natural numbers.
Note that \texttt{Lam} is actually just a nested datatype 
\cite{alti:csl99} while \texttt{Fin} exploits the full power of
GADTs because the range of the constructors is constrained.

\todo{Rewrite and expand}

The paper is an expanded and revised version of the LICS paper by the
first and 3rd author \cite{lics}. In the present paper we have
integrated the Agda formalisation in the main development, which in
many instances required extending it. We have made explicit the use of
relative monads which was only hinted at in the conference version
based on the recent work on relative monads \cite{relmon}. We have
also dualized the development to terminal coalgebras which requires a
non-trivial change in the setup (section \ref{sec:termcoalg}).  We
have also formalized the derivation of indexed W-types from ordinary
W-types (section \ref{wifromw}. The derivation of M-types from W-types
(section \ref{sec:mfromw})
was already given in \cite{C-CSPTs} is revisited here exploiting the
indexed W-type derived previously amd the development is formalized in
Agda.

@@

Our contructions are all developed in Agda, and so we adopt its syntax, but we will take certain liberties with its type-theory.

@@

\noindent
As a notational convenience, we will continue to define extra Agda data-types in the rest of the paper, but in the end we will show how each of these can be reduced to a theory that contains only \ensuremath{\Conid{W}}. For compactness, and readablity we will also define functions using Agda's pattern matching syntax, rather than encoding them using wrec, it is an unstated lemma that each of these definitions can be reduced to terms which only use \ensuremath{\Varid{wrec}}.

@@

This is an intensioanl equality, but we want to work in a setting with extensional type-theory, so we extend the propositional equality with this extensionality axiom:

@@

This creates non-canonical elements of \ensuremath{\Varid{\char95 ≅\char95 }}, \emph{i.e.} closed terms in equality types which are not \ensuremath{\Varid{refl}}. In order to deal with these non-canonical elements, we also rely on axiom \ensuremath{\Conid{K}}, or the uniqueness of idenity proofs:

@@

@@

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
\column{28}{@{}>{\hspre}l<{\hspost}@{}}%
\column{69}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{record}\;\Conid{Cont}\;\in\;\Conid{Set₁}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{constructor}\;\anonymous \lhd\anonymous {}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{field}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Conid{S}\;\in\;\Conid{Set}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\Conid{Cont}\;\rightarrow\;\Conid{Func}{}\<[E]%
\\
\>[B]{}\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\mathrel{=}\;\Keyword{record}\;{}\<[21]%
\>[21]{}\{\mskip1.5mu \Varid{obj}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Conid{A}\;\rightarrow\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\!\times\!\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{A}){}\<[E]%
\\
\>[21]{};\Varid{mor}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\ensuremath{\mbox{$\circ$}}\;\Varid{f})\;{}\<[69]%
\>[69]{}\!\;\!{}\<[E]%
\\
\>[21]{}\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

@@

The category of containers is a full and faithful sub-category of the functor category. We have also shown that the category of containers is cartesian closed, and is closed under formation of co-products.

@@

However, we have also shown that for \ensuremath{\Varid{n}}-ary containers (containers with \ensuremath{\Varid{n}} position sets), it is possible to define a \emph{parameterised} initial algebra construction \ensuremath{\mu\;\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;\rightarrow\;\Conid{Cont}\;(\Varid{suc}\;\Varid{n})\;\rightarrow\;\Conid{Cont}\;\Varid{n}}. This allows us to model a broad range of nested and mutual types as containers. We will develop an analogous construction for indexed conatainers in section ???.

@@

\noindent
such that both \ensuremath{\Varid{id}^{\star}} is mapped to \ensuremath{\Varid{id}} and \ensuremath{\anonymous \circ^{\star}\anonymous } to \ensuremath{\anonymous \ensuremath{\mbox{$\circ$}}\anonymous } under the action of 
\ensuremath{\Varid{mor}}. We adopt the convention that the projections \ensuremath{\Varid{obj}} and \ensuremath{\Varid{mor}} are silent, 
\emph{i.e.} depending on the context \ensuremath{\Conid{F}\;\in\;\Conid{IFunc}\;\Conid{I}} can stand for either the 
functor's action on objects, or on morphisms. A morphism between to such 
indexed functors is a natural transormation:

@@

\noindent
This construction is used, for instance, in building the pattern functor for \ensuremath{\Conid{ScLam}} as in the introduction; Concentranting only on the \ensuremath{\Varid{abs}} case we want to build  
\ensuremath{\Conid{ScLam′}\;\Conid{X}\;\Varid{n}\;\mathrel{=}\;(\Conid{X}\;\ensuremath{\mbox{$\circ$}}\;\Varid{suc})\;\Varid{n}}. Or simply \ensuremath{\Conid{ScLam′}\;\Conid{X}\;\mathrel{=}\;\Delta^{\text{\tiny F}}\;\Varid{suc}\;\Conid{X}}. In general this combinator 
restricts the functor \ensuremath{\Conid{X}} to the indicies in the image of the
function \ensuremath{\Varid{f}}.

@@

\noindent
Perhaps unsuprisingly, \ensuremath{\Sigma^{\text{\tiny F}}} turns out to be the left adjoint to re-indexing 
(\ensuremath{\Delta^{\text{\tiny F}}}). Its right adjoint, we denote \ensuremath{\Pi^{\text{\tiny F}}}:

@@

For the morphism part of an indexed-functor over a co-product we can eliminate
the coproduct and curry the result ing definition in this way:

@@

Which follow from the extensionality of our propositional equality, the 
assosciativity of \ensuremath{\Sigma} and the terminality of \ensuremath{\top}. By the full and faithful 
nature of the embedding into \ensuremath{\Conid{IFunc}} we can then reuse the result that
\ensuremath{(\Conid{IFunc},\eta^{\text{\tiny F}},\anonymous \bind ^{\text{\tiny F}}\anonymous )} is a relative monad.

@@

Again we appeal to the full and faithfullness of \ensuremath{\Varid{⟦\char95 ⟧\char95 }} and show instead that
the embedding into \ensuremath{\Conid{IFunc}} preserves these constructions. That, is we want to show that the following three statements hold:

@@

\noindent
The composite container has shapes given by a shape \ensuremath{\Varid{s}\;\in\;\Conid{S}} and an assigment of \ensuremath{\Conid{T}} 
shapes to \ensuremath{\Conid{P}^{\Conid{J}}\;\Varid{s}} positions. Positions are then a choice between 
a \ensuremath{\Conid{I}}-indexed position, or a pair of an \ensuremath{\Conid{J}}-indexed position, and a \ensuremath{\Conid{Q}} position
\emph{underneath} the appropriate \ensuremath{\Conid{T}} shape.

@@

We will now examine how to construct the initial algebra of a container of the form \ensuremath{\Conid{F}\;\in\;\Conid{ICont}^{\star}\;(\Conid{I}\;\uplus\;\Conid{J})\;\Conid{J}}. The shapes of such a container are an \ensuremath{\Conid{J}}-indexed family of \ensuremath{\Conid{Set}}s and the positions are indexed by \ensuremath{\Conid{I}\;\uplus\;\Conid{J}}; we will treat these position as two separate entities, those positions indexed by \ensuremath{\Conid{I}} -- the recusive positions -- and those by \ensuremath{\Conid{J}} -- the payload positions.

@@

\noindent
We can now give the object part of the patrametrized initial algebra of a container, given by:

@@

\noindent
The algebra map is a container morphism from the partial aplication of \ensuremath{\Conid{F}} and its parametrised initial algebra, to the initial algebra, given by the algebra map of \ensuremath{\Conid{WI}} (\ensuremath{\Varid{sup}}) and our mediation funtion \ensuremath{\Varid{path}}:

@@

Dually to the initial algebra construction out-lined above, we can also show that indexed-containers are closed under parameterized terminal co-algebras. We proceed in much the same way as before, by first constructing the dual of the indexed \ensuremath{\Conid{W}}-type, which we refer to as an indexed \ensuremath{\Conid{M}}-type. As you might expect this is in fact the plain (as opposed to parametrized) terminal co-algebra of an indexed container functor:

@@

@@

How, then, can we build \ensuremath{\Conid{WI}} from \ensuremath{\Conid{W}}? The initial step is to create a type of \emph{pre}-\ensuremath{\Conid{WI}} trees, with nodes containing a shape \emph{and} its index, and branching over positions \emph{and their} indicies:

@@

\noindent
We can truncate any given tree of depth greater than 1, this amounts to the 
iteration of the morphism part of the container functor applied to the unique 
moprhism into the terminal object:

@@

So now we can build the chain of finitie itererations of a container functor:

@@

\begin{proof}
To prove the structure is a relative 
monad we observe that the following equlities hold:

>>>>

To model this we move to the notion of an inductive
family in Type Theory. A family is a type indexed by another already
given type. The first example is the family of finite sets \ensuremath{\Conid{Fin}} which
assigns to any natural number \ensuremath{\Varid{n}} a set \ensuremath{\Conid{Fin}\;\Varid{n}} which has exactly
\ensuremath{\Varid{n}} elements. \ensuremath{\Conid{Fin}} can be used where in conventional reasoning we
assume any finite set, e.g. when dealing with a finite address apace or
a finite set of variables. The inductive definition of \ensuremath{\Conid{Fin}} refines
the type of natural numbers:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\Conid{Fin}\;\in\;\Conid{ℕ}\;\rightarrow\;\Conid{Set}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{zero}\;{}\<[9]%
\>[9]{}\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;{}\<[30]%
\>[30]{}\rightarrow\;\Conid{Fin}\;(\Varid{suc}\;\Varid{n}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{suc}\;{}\<[9]%
\>[9]{}\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;(\Varid{i}\;\in\;\Conid{Fin}\;\Varid{n})\;{}\<[30]%
\>[30]{}\rightarrow\;\Conid{Fin}\;(\Varid{suc}\;\Varid{n}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

\noindent
Inductive families are the backbone of
dependently typed programming as present in Epigram or
Agda~\cite{Agda}. Coq also supports the definition of inductive families
but programming with them is rather hard --- a situation which has been
improved by the new \texttt{Program} tactic~\cite{sozeau}. 
More recently, the implementation of Generalized Algebraic Datatypes 
(GADTs)~\cite{Hinze:GADT} 
allows \ensuremath{\Conid{Fin}} and \ensuremath{\Conid{Lam}} to be encoded in Haskell:
\begin{tabbing}\tt
~data~Fin~a~where~\\
\tt ~~~FZero~\char58{}\char58{}~Fin~\char40{}Maybe~a\char41{}\\
\tt ~~~FSucc~\char58{}\char58{}~Fin~a~\char45{}\char62{}~Fin~\char40{}Maybe~a\char41{}\\
\tt ~\\
\tt ~data~Lam~a~where~\\
\tt ~~~Var~\char58{}\char58{}~Fin~a~\char45{}\char62{}~Lam~a\\
\tt ~~~App~\char58{}\char58{}~Lam~a~\char45{}\char62{}~Lam~a~\char45{}\char62{}~Lam~a\\
\tt ~~~Abs~\char58{}\char58{}~Lam~\char40{}Maybe~a\char41{}~\char45{}\char62{}~Lam~a
\end{tabbing}
Here \texttt{Fin} and \texttt{Lam} are indexed by types instead of
natural numbers; The type constructor \texttt{Maybe} serves as a type level
copy of the \ensuremath{\Varid{succ}} constructor for natural numbers.
Note that \texttt{Lam} is actually just a nested datatype 
\cite{alti:csl99} while \texttt{Fin} exploits the full power of
GADTs because the range of the constructors is constrained.

@@

\noindent We introduce the notion of indexed container which
generalizes containers allowing us to represent inductive
families. This is a further step from \emph{dependent polynomial
  functors} \cite{HylandGambino} representing endofunctors on a slice
category. Indexed containers as introduced in the present paper allow us
to represent functors between different slices and capture also mutual
and nested inductive definitions.

While Hyland and Gambino \cite{HylandGambino} show that dependent polynomial 
functors always
have initial algebras, we show that indexed containers are closed under 
parametrized initial algebras. Hence we can apply the fixpoint
construction several times. The flexibility of indexed
containers allows us to also establish closure under the adjoints of
reindexing. This leads directly to a grammar for strictly positive
families, which itself is an instance of a strictly positive family 
(section \ref{sec:spf}) --- see also our previous work \cite{alti:cats07,alti:jcats07}.

Containers are related to Girard's normal functors \cite{GirardNormal} which
themselves are a special case of Joyal's analytic functors
\cite{JoyalA:fonaes} --- those that allow only finite sets of positions.
Fiore, Gambino, Hyland and Winskel's work on generalized species
\cite{fiore2008ccb} considers those concepts in a more generic setting ---
th e precise relation of this work to indexed containers remains to be
explored.

Perhaps the earliest publication related to indexed containers
occurs in Petersson and Synek's paper
\cite{PS89} from 1989. They present rules extending Martin-L{\"o}f's
type theory with a set constructor for `tree sets' : families of
mutually defined inductive sets, over a fixed index set.

Inspired in part by Petersson and Synek's constructor,
Hancock, Hyvernat and Setzer \cite{hancock-apal06} applied indexed (and unindexed)
containers, under the name `interaction structures' to the task of
modelling imperative interfaces such as command-response interfaces in
a number of publications.

This paper is an expanded and revised version of the LICS paper by the
first and 3rd author \cite{lics}. In the present paper we have
integrated the Agda formalisation in the main development, which in
many instances required extending it. We have made explicit the use of
relative monads which was only hinted at in the conference version
based on the recent work on relative monads \cite{relmon}. We have
also dualized the development to terminal coalgebras which required
the type of paths to be defined inductively instead of recursively as
done in the conference paper (section \ref{sec:termcoalg}).  We
have also formalized the derivation of indexed W-types from ordinary
W-types (section \ref{wifromw}. The derivation of M-types from W-types
(section \ref{sec:mfromw})
was already given in \cite{C-CSPTs} is revisited here exploiting the
indexed W-type derived previously. moreover the development is formalized in
Agda.

@@

Our constructions are all developed in Agda, and so we adopt its syntax, but we will take certain liberties with its type-theory.

@@

\noindent
As a notational convenience, we will continue to define extra Agda data-types
in the rest of the paper, but in the end we will show how each of these can
be reduced to a theory that contains only \ensuremath{\Conid{W}}. For compactness, and
readability we will also define functions using Agda's pattern matching
syntax, rather than encoding them using \ensuremath{\Varid{wrec}}, it is an unstated lemma that
each of these definitions can be reduced to terms which only use \ensuremath{\Varid{wrec}}.

@@

This is an intensional equality, but we want to work in a setting with extensional type-theory, so we extend the propositional equality with this extensionality axiom:

@@

This creates non-canonical elements of \ensuremath{\Varid{\char95 ≅\char95 }}, \emph{i.e.} closed terms in
equality types which are not \ensuremath{\Varid{refl}}. In order to deal with these
non-canonical elements, we also rely on axiom \ensuremath{\Conid{K}}, or the uniqueness of
identity proofs:

@@

With these assumptions we obtain a theory which corresponds to
extensional Type Theory \cite{hofmann-conservativity}.

@@

The initial algebra semantics is useful to provide a generic
analysis of inductive types exploiting generic concepts such as
constructors and functorial map. However, it cannot say whether such inductive
types actually exist, and it falls short of providing a systematic
characterisation of generic operations such as equality or the
zipper~\cite{huet:zipper,conor:derivative}.

In previous
work, \cite{alti:cont-tcs,alti:fossacs03}, we have proposed the notion of a container
type: A (unary) container is given by a set of shapes \ensuremath{\Conid{S}} and a
family of positions \ensuremath{\Conid{P}} assigning, to each shape, the set
of positions where data can be stored in a data structure of that
shape.

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{record}\;\Conid{Cont}\;\in\;\Conid{Set₁}\;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{constructor}\;\anonymous \lhd\anonymous {}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Keyword{field}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Conid{S}\;\in\;\Conid{Set}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Conid{P}\;\in\;\Conid{S}\;\rightarrow\;\Conid{Set}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

Every container \ensuremath{\Conid{S}\;\lhd\;\Conid{P}}  gives rise to a functor:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
\column{28}{@{}>{\hspre}l<{\hspost}@{}}%
\column{69}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{⟦\char95 ⟧}\;\in\;\Conid{Cont}\;\rightarrow\;\Conid{Func}{}\<[E]%
\\
\>[B]{}\Varid{⟦}\;\Conid{S}\;\lhd\;\Conid{P}\;\Varid{⟧}\;\mathrel{=}\;\Keyword{record}\;{}\<[21]%
\>[21]{}\{\mskip1.5mu \Varid{obj}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Conid{A}\;\rightarrow\;\Sigma(\!\;\Varid{s}\;\in\;\Conid{S}\;\!)\!\times\!\;(\Conid{P}\;\Varid{s}\;\rightarrow\;\Conid{A}){}\<[E]%
\\
\>[21]{};\Varid{mor}\;{}\<[28]%
\>[28]{}\mathrel{=}\;\lambda\;\Varid{m}\;\rightarrow\;\lambda (\!\;\Varid{s}\;\!,\!\;\Varid{f}\;\!\!)\;\rightarrow\;(\Varid{s},\Varid{m}\;\ensuremath{\mbox{$\circ$}}\;\Varid{f})\;{}\<[69]%
\>[69]{}\!\;\!{}\<[E]%
\\
\>[21]{}\mskip1.5mu\}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

We define morphisms of containers, this can be derived from the specification that embedding should be full and faithful using the Yoneda lemma:

@@

The category of containers is a full and faithful sub-category of the functor category. We have also shown that the category of containers is cartesian closed (\cite{cie}, and is closed under formation of co-products.

@@

However, we have also shown that for \ensuremath{\Varid{n}}-ary containers (containers with \ensuremath{\Varid{n}} position sets), it is possible to define a \emph{parameterised} initial algebra construction \ensuremath{\mu\;\in\;\forall{}\;\{\mskip1.5mu \Varid{n}\mskip1.5mu\}\;\rightarrow\;\Conid{Cont}\;(\Varid{suc}\;\Varid{n})\;\rightarrow\;\Conid{Cont}\;\Varid{n}}. This allows us to model a broad range of nested and mutual types as containers.

@@

\noindent
such that both \ensuremath{\Varid{id}^{\star}} is mapped to \ensuremath{\Varid{id}} and \ensuremath{\anonymous \circ^{\star}\anonymous } to \ensuremath{\anonymous \ensuremath{\mbox{$\circ$}}\anonymous } under the action of 
\ensuremath{\Varid{mor}}. We adopt the convention that the projections \ensuremath{\Varid{obj}} and \ensuremath{\Varid{mor}} are silent, 
\emph{i.e.} depending on the context \ensuremath{\Conid{F}\;\in\;\Conid{IFunc}\;\Conid{I}} can stand for either the 
functor's action on objects, or on morphisms. A morphism between to such 
indexed functors is a natural transformation:

@@

\noindent
This construction is used, for instance, in building the pattern functor for
\ensuremath{\Conid{ScLam}} as in the introduction; Concentrating only on the \ensuremath{\Varid{abs}} case we want
to build  \ensuremath{\Conid{ScLam′}\;\Conid{X}\;\Varid{n}\;\mathrel{=}\;(\Conid{X}\;\ensuremath{\mbox{$\circ$}}\;\Varid{suc})\;\Varid{n}}. Or simply \ensuremath{\Conid{ScLam′}\;\Conid{X}\;\mathrel{=}\;\Delta^{\text{\tiny F}}\;\Varid{suc}\;\Conid{X}}. In
general this combinator restricts the functor \ensuremath{\Conid{X}} to the indices in the
image of the function \ensuremath{\Varid{f}}.

@@

\noindent
Perhaps unsurprisingly, \ensuremath{\Sigma^{\text{\tiny F}}} turns out to be the left adjoint to re-indexing 
(\ensuremath{\Delta^{\text{\tiny F}}}). Its right adjoint, we denote \ensuremath{\Pi^{\text{\tiny F}}}:

@@

For the morphism part of an indexed-functor over a co-product we can eliminate
the co-product and curry the resulting definition in this way:

@@

Which follow from the extensionality of our propositional equality, the 
associativity of \ensuremath{\Sigma} and the terminality of \ensuremath{\top}. By the full and faithful 
nature of the embedding into \ensuremath{\Conid{IFunc}} we can then reuse the result that
\ensuremath{(\Conid{IFunc},\eta^{\text{\tiny F}},\anonymous \bind ^{\text{\tiny F}}\anonymous )} is a relative monad.

@@

Again we appeal to the full and faithfulness of \ensuremath{\Varid{⟦\char95 ⟧\char95 }} and show instead that
the embedding into \ensuremath{\Conid{IFunc}} preserves these constructions. That, is we want to show that the following three statements hold:

@@

\noindent
The composite container has shapes given by a shape \ensuremath{\Varid{s}\;\in\;\Conid{S}} and an assignment
of \ensuremath{\Conid{T}} shapes to \ensuremath{\Conid{P}^{\Conid{J}}\;\Varid{s}} positions. Positions are then a choice between a
\ensuremath{\Conid{I}}-indexed position, or a pair of an \ensuremath{\Conid{J}}-indexed position, and a \ensuremath{\Conid{Q}}
position \emph{underneath} the appropriate \ensuremath{\Conid{T}} shape.

@@

We will now examine how to construct the initial algebra of a container of
the form \ensuremath{\Conid{F}\;\in\;\Conid{ICont}^{\star}\;(\Conid{I}\;\uplus\;\Conid{J})\;\Conid{J}}. The shapes of such a container are an
\ensuremath{\Conid{J}}-indexed family of \ensuremath{\Conid{Set}}s and the positions are indexed by \ensuremath{\Conid{I}\;\uplus\;\Conid{J}}; we
will treat these position as two separate entities, those positions indexed
by \ensuremath{\Conid{I}} -- the recursive positions -- and those by \ensuremath{\Conid{J}} -- the payload
positions.

@@

\noindent
We can now give the object part of the parametrised initial algebra of a
container, given by:

@@

\noindent
The algebra map is a container morphism from the partial application of \ensuremath{\Conid{F}}
and its parametrised initial algebra, to the initial algebra, given by the
algebra map of \ensuremath{\Conid{WI}} (\ensuremath{\Varid{sup}}) and our mediation function \ensuremath{\Varid{path}}:

@@

Dually to the initial algebra construction out-lined above, we can also show
that indexed-containers are closed under parameterised terminal co-algebras.
We proceed in much the same way as before, by first constructing the dual of
the indexed \ensuremath{\Conid{W}}-type, which we refer to as an indexed \ensuremath{\Conid{M}}-type. As you might
expect this is in fact the plain (as opposed to parametrized) terminal
co-algebra of an indexed container functor:

@@

So far we have developed a theory of indexed containers using a rich Type
Theory. We claimed in the introduction, however, that this theory required
only the presence of \ensuremath{\Conid{W}}-types. In this section we will outline the
translation of many of the definitions above into such a spartan theory.
First we will show how to obtain indexed \ensuremath{\Conid{W}}-types from un-indexed ones, and
by analogy \ensuremath{\Conid{MI}} from \ensuremath{\Conid{M}}, and then we will revisit a proof of how to derive
\ensuremath{\Conid{M}}-types from \ensuremath{\Conid{W}}.

@@

How, then, can we build \ensuremath{\Conid{WI}} from \ensuremath{\Conid{W}}? The initial step is to create a type
of \emph{pre}-\ensuremath{\Conid{WI}} trees, with nodes containing a shape \emph{and} its index,
and branching over positions \emph{and their} indices:

@@

\noindent
We can truncate any given tree of depth greater than 1, this amounts to the 
iteration of the morphism part of the container functor applied to the unique 
morphism into the terminal object:

@@

So now we can build the chain of finite iterations of a container functor:

@@

\begin{proof}
To prove the structure is a relative 
monad we observe that the following equalities hold:
